Species persistence under climate and fishing

### 2014-01-23 Writing/MS.tex
Working on intro revisions, finished page 1 of Malin edits for #1

There are many stressors that can disturb an ecosystem. Ecologists have quantified the effects of a number of stressors individually \citep{Wilcoveetal1998, Crainetal2008, DarlingCote2008}. However, disturbances rarely occur in isolation.  It is therefore important to understand how a system is affected by multiple disturbances together \citep{DoakMorris2010, Fordhametal2013, Foltetal1999}. A perturbation that has little effect when affecting a system individually may amplify the disturbance caused by a coincident perturbation \citep{Crainetal2008, DarlingCote2008}. Synergistic interactions between stressors are especially worrying if an animal population can persist in the face of a single stressor but not in the presence of multiple stressors (e.g.  \citet{Pelletieretal2006}). However, it remains hard to predict when disturbances are likely to interact in ways not predicted by the effects of individual perturbations.

Climate change is a prime example of a stressor occurring in a background of multiple other, possibly interacting, disturbances. These include pollution, habitat fragmentation, invasive species, and over-exploitation \citep{Wilcoveetal1998, Salaetal2000, MEA2005}. It is expected that  species will move their home ranges in response to climate change. Many marine fish are already moving \citep{Perryetal2005, HiddinkHoftstede2008, Rijnsdorpetal2009, Dulvyetal2008, Simpsonetal2011} and are projected to continue to move in the future \citep{Kelletal2005, Mackenzieetal2007}. These species are faced not only with temperature-driven range range shifts but also with anthropogenic stressors including ocean acidification and commercial fishing \citep{Pinskyetal2013, Barryetal1995, Nyeetal2009}. Climate change and fishing have been identified as the two largest human impacts on the ocean \citep{Halpernetal2008} so there is a need to understand whether and how these disturbances will interact in the future.

Many methods have been proposed to help mitigate the effects of harvesting on fish populations. Marine protected areas (MPAs), which provide stepping stones to help species keep up with a changing environment, have been suggested as a key form of climate insurance \citep{Thomasetal2012, Hannahetal2007}. MPAs are frequently recommended for both conservation of biodiversity and improved fisheries yield \citep{Gainesetal2010}. Threshold harvesting levels below which no fish may be harvested have also been proposed as a means of both optimizing yield and conserving resources \citep{Landeetal1995, Landeetal1997}.

A wide range of models have been used to study the effect of climate
change and harvesting on fish populations.  A common approach to predicting future population distributions under climate change has been to use bioclimatic-envelope models (also known as species distribution models -- SDMs). These statistical models have been
used to find correlations between the presence or absence of fish
species and biophysical properties of the environment such as mean or maximum
temperatures, rainfall, or salinity, to explain and predict how species
ranges will differ under climate change \citep{Elithetal2006, GuisanThuiller2005, GuisanZimmerman2000}. Similarly,
projection models of future climate conditions have been used to predict
where conditions will be favorable for fish populations in the future
\citep{Cheungetal2008, Cheungetal2009}.  Despite these models' widespread
adoption, they have  been criticized for being oversimplified
as they lack species interactions, dispersal and reproductive processes
\citep{KearneyPorter2009, Zarnetskeetal2012, Robinsonetal2011}.

>>>>

There are many stressors that can disturb an ecosystem, ecologists have quantified the effects of a number of stressors individually \citep{Wilcoveetal1998, Crainetal2008, DarlingCote2008}. However, disturbances rarely occur in isolation.  It is therefore important to understand how a system is affected by multiple disturbances together \citep{DoakMorris2010, Fordhametal2013, Foltetal1999}. A perturbation that has little effect when affecting a system may amplify the disturbance caused by a coincident perturbation \citep{Crainetal2008, DarlingCote2008}. Synergistic interactions between stressors are especially worrying if an animal population can persist in the face of a single stressor but not in the presence of multiple stressors (e.g.  \citet{Pelletieretal2006}). However, it remains difficult to predict when disturbances are likely to interact in ways not anticipated by the effects of individual perturbations.

Climate change is a prime example of a stressor occurring in a background of multiple other, possibly interacting, disturbances such as pollution, habitat fragmentation, invasive species, and over-exploitation \citep{Wilcoveetal1998, Salaetal2000, MEA2005}. Climate change is expected to drive range shifts;  many marine fish, for example, are already moving \citep{Perryetal2005, HiddinkHoftstede2008, Rijnsdorpetal2009, Dulvyetal2008, Simpsonetal2011} and are projected to continue to move in the future \citep{Kelletal2005, Mackenzieetal2007}. These species are faced not only with temperature-driven range range shifts but also with anthropogenic stressors including ocean acidification and commercial fishing \citep{Pinskyetal2013, Barryetal1995, Nyeetal2009}. Climate change and fishing have been identified as the two largest human impacts on the ocean \citep{Halpernetal2008} so there is a need to understand whether and how these disturbances will interact.

Many management tactics have been proposed to help mitigate the effects of harvesting on fish populations. Marine protected areas (MPAs), which provide stepping stones to help species keep up with a changing environment, have been suggested as a key form of climate insurance \citep{Thomasetal2012, Hannahetal2007}. MPAs are frequently recommended for both conservation of biodiversity and improved fisheries yield \citep{Gainesetal2010}. Threshold harvesting levels below which no fish may be harvested have also been proposed as a means of both optimizing yield and conserving resources \citep{Landeetal1995, Landeetal1997}.

A wide range of models have been used to study the effect of climate
change and harvesting on fish populations.  A common approach to predicting future population distributions under climate change has been to use bioclimatic-envelope models (also known as species distribution models -- SDMs). In terrestrial systems these statistical models have been
used to find correlations between the presence or absence of fish
species and biophysical properties of the environment such as mean or maximum
temperatures, rainfall, or salinity, to explain and predict how species
ranges will differ under climate change \citep{Elithetal2006, GuisanThuiller2005, GuisanZimmerman2000}. Similarly,
projection models of future climate conditions have been used to predict
where conditions will be favorable for fish populations in the future
\citep{Cheungetal2008, Cheungetal2009, Hareetal2010, Lenoiretal2011}.  Despite these models' widespread
adoption, they have  been criticized for being oversimplified
as they lack species interactions, dispersal and reproductive processes
\citep{KearneyPorter2009, Zarnetskeetal2012, Robinsonetal2011}.
